A Comparative Ecophysiological Survey of Lycophytes, Ferns, and Seed Plants

### 2019-02-18 manuscript/manuscript.Rmd
started stats for photo and gs

We examined a series of fern, lycophyte, and angiosperm species at both sites (Table 1). The main focus of this work was to examine ferns relative to light-analogous angiosperms. To accomplish this, we selected a fern species and then located the nearest neighboring angiosperm growing in similar habitat conditions. Surveyed angiosperm taxa were always herbaceous and similar in size and stature to the neighboring fern species. At the ENHP, we sampled across two distinct habitats: understory low-light areas and open, high-light areas. At the Colgate University site ......

@@

For the light response curves, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm). Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels. These measurements were used to generate light response curves using appropriate light levels for high-light and low-light species separately. The light levels were not identical for each species, with several points being added or subtracted with each run. Broadly, light response curves followed the same template for high light or low light species. For high light species, light levels started at a PPFD of 1500 $\mu$mol m^-1^ s^-1^ and then consisted of ~16 additional steps to O $\mu$mol m^-1^ s^-1^. For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^. From these light response curves, we developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates. As a result, additional survey measurments were conducted at at a PPFD of 1200 $\mu$mol m^-1^ s^-1^ for high-light species and 600 $\mu$mol m^-1^ s^-1^ for low-light species. All other cuvette parameters were identical to those used for the light response curves.

@@

Air-dried and ground tissues from each leaf sample used for gas exchange were analyzed for total C and N. Foliar carbon and nitrogen content were generated using a costech elemental analyzer (machine). Phosphorus data were gathered using some digest (method)

@@

@@

@@

@@

Across closed canopy environments where lycophytes, some of these differences between angiosperms and ferns break down. For example, there is no significant difference between the photosynthetic rates of angiosperms and ferns, although these two groups are significantly different than lycophytes. This relationship is also shown in internal CO2 concentrations, with angiosperms and ferns showing no significant difference and lycophytes differing significantly from both groups. For stomatal conductances, however, all three groups are significantly different (Figure adfd).

# Nutrient relations

>>>>

We examined a series of fern, lycophyte, and angiosperm species at both sites (Table 1). The main focus of this work was to examine ferns relative to light-analogous angiosperms. To accomplish this, we selected a fern species and then located the nearest neighboring angiosperm growing in similar habitat conditions. Surveyed angiosperm taxa were always herbaceous and similar in size and stature to the neighboring fern species. At the ENHP, we sampled across two distinct habitats: understory low-light areas and open, high-light areas. At the Colgate University site, additioanl low-light fern and lycophyte species were surveyed.

@@

Light response curves were conducted by decreasing the light intensity in the gas exchange cuvette in small step changes. For each light response curve, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm). Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels. These measurements were used to generate light response curves using appropriate light levels for high-light and low-light species separately. The light levels were not identical for each species, with several points being added or subtracted with each run. Broadly, light response curves followed the same template for high light or low light species. For high light species, light levels started at a PPFD of 1500 $\mu$mol m^-1^ s^-1^ and then consisted of ~16 additional steps to O $\mu$mol m^-1^ s^-1^. For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^. The light compensation point (LCP) was calculated as the PPFD at which the net photosynthetic rate equalled zero from the linear phase of each light response curve. Quantum yeild (phi) was calculated as the initial slope of the assimilation rate and dark respiration rate was recorded at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.

From these light response curves, we developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates. As a result, additional survey measurments were conducted at at a PPFD of 1200 $\mu$mol m^-1^ s^-1^ for high-light species and 600 $\mu$mol m^-1^ s^-1^ for low-light species. All other cuvette parameters were identical to those used for the light response curves.

@@

Foliar tissue was sampled following gas exchange measurements for all species at ENHP and used for nutrient analyses. Samples were dried to a constant mass and ground using a Wig-L-Bug (Sigma-Aldrich Co. St. Louis, USA). Carbon and nitrogen analyses were measured using a Costech Analytical Elemental Analyzer (Valencia, USA), with the percentage of carbon and nitrogen in samples calculated by comparison with certified standards. Foliar phosphorus concentrations were determined using an ash digestion process (D’Angelo et al., 2001) preceded by colour development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon).

Photosynthetic nitrogen use efficiency (PNUE) was defined as the ratio of photosynthesis to leaf nitrogen content on a mass basis.*

@@

Stomatal density (SD) was measured by directly counting stomata on the abaxial leaf surface under 40x magnification with a field of view of 0.622 mm^2. Stomatal density was calculated from 8 non-overlapping foliar regions for five individuals of each species.

@@

Linear mixed-effect models responses were used to test responses of functional traits to caetgorial fixed effects of plant group and canopy habitat. To test broad differnces in plant groups, species nested within canopy habitats were treated as random effects. To test differences between habitat types (ferns and angiosperms only) species was set as a fixed effect. Explained variance (R2) of mixed models were computed as in Nakagawa and Schielzeth (2013), in which the marginal R2 represents variance explained by fixed factors and the conditional R2 by both fixed and random factors.Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package (Hothorn et al., 2008). All tests of statistical significance were conducted at an α level of 0.05. All analyses were performed with R 3.5.1 [R cite].

T-tests were performed to test for differences between functional traits, pooled for species, across canopy habitats within angioperm and fern plant groups.

@@

Overall, fern and angiosperm species in open habitats had higher A~n~ than compaable species in low-light habitats (*P* <- 0.0001). Specifically, fern species in high-light habitats had 40.6 % greater A~n~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 60 % higher A~n~ (*P* < 0.0001, t=-6.97).
Ferns and angiosperms growing in low-light habitats had similar A~n~, however, in open habitats angiosperms had 42.6% higher rates of A~n~ than ferns (canopy x plant group, *P* = 0.043).

Specifically, fern species in high-light habitats had 50.7 % greater g~s~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 66.9 % higher g~s~ (*P* < 0.0001, t=-4.157).

@@

Across closed canopy environments, some of these differences between angiosperms and ferns break down when the model includes lycophytes. No differences between A~n~ were detected between between plant groups, with a majority of the variation associated with the random effect of individual species. Stomatal conducance was higher in angiosperms than lycophytes, while ferns were not different from either plant group (*P* = 0.005). The leaf internal CO~2~ concentrations were not different between plant groups, with a majority of the variation associated with the random effect of individual species.

# Stoichiometry
